Fast Hyperparameter Selection for Graph Kernels via Subsampling and\\Multiple Kernel Learning

### 2016-11-26 paper.tex
F

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
%\author{First author$^1$ and Second author$^2$
\author{Michele Donini$^1$, Nicol\`o Navarin$^2$, Ivano Lauriola$^2$, Fabio Aiolli$^2$ and Fabrizio Costa$^3$
%
% Optional short acknowledgment: remove next line if non-needed
%\thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- Computational Statistics and Machine Learning (CSML) \\
Istituto Italiano di Tecnologia, Via Morego 30, Genova, Italy
\vspace{.1cm}\\
2- Department of Mathematics, University of Padova\\
via Trieste 63, Padova, Italy
\vspace{.1cm}\\
3- ... \\
...
}

@@

\begin{abstract}
Kernels for structures are generally characterised by a number of hyperparameters which need to be tuned. Moreover, different types of kernels can be used and the best choice usually depends on data. The hyperparameter setting is a time-consuming step. One option proposed to overcome this problem is to use the Multiple Kernel Learning approach and generate a kernel which is a combination of different types of kernels and different hyperparameters settings. However, this solution still requires the computation of many large kernel matrices. In this paper we propose a subsampling method to efficiently select just a small number of kernels which need to be actually computed. Finally, we give empirical evidence that the method proposed is far more efficient than the other approaches while maintaining similar results in terms of AUC.
\end{abstract}

\section{Introduction}
In different real-world domains, it may be easy to encode examples in a structured form. For example, in chemoinformatics there is a well known graph representation of the secondary structures of chemical compounds. Moreover, images can easily be represented as graph after segmentation, where adjacent objects (or regions) are linked.
In these cases, deriving a vectorial representation for the data may be a time-consuming task, where information from  domain experts is needed, and information loss is, in general, unavoidable.
Another option is to deal directly with examples represented in a structured form.
In this case, specific machine learning techniques have to be adopted.
Kernel methods are one option, where it suffices to adopt a kernel function defined for data type under consideration. Then, standard (kernelized) learning algorithms (e.g. SVM) can be adopted.

>>>>

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
%\author{First author$^1$ and Second author$^2$
\author{Michele Donini$^1$, Nicol\`o Navarin$^2$, Ivano Lauriola$^2$, Fabio Aiolli$^2$ and Fabrizio Costa$^3$
%
% Optional short acknowledgment: remove next line if non-needed
%\thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- Computational Statistics and Machine Learning (CSML) \\
Istituto Italiano di Tecnologia, Via Morego 30, Genova, Italy
\vspace{.1cm}\\
2- Department of Mathematics, University of Padova\\
via Trieste 63, Padova, Italy
\vspace{.1cm}\\
3- Department of Computer Science, Freiburg University \\
Georges-Kohler-Allee 106, 79110 Freiburg, Germany\\
}

@@

\begin{abstract}

Model selection is one of the most computationally expensive tasks in a
machine learning application. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned.  Multiple Kernel Learning offers a way to approach this
computational bottleneck by generating a combination of different kernels
under different parametric settings. However, this solution still requires the
computation of many large kernel matrices. In this paper we propose a method
to efficiently select just a small number of kernels without loss in
predictive performance.

\end{abstract}

\section{Introduction}

Many application domains can be naturally encoded in a structured form. In
chemo- and bioinformatics for example, molecules and polymers can be encoded
as graphs with atoms as nodes and chemical bonds as edges. After a
segmentation pre-processing step, images can also be represented as networks
of adjacent objects or regions. Working on these domains employing traditional
feature engineering is possible, but it is generally time-consuming (every
different application requires a novel feature design by some expert) and
information loss is often unavoidable. Direct processing of structures is
possible using a kernelized approach, where a kernel function can be devised
for each generic data type (for example sequences, trees or graphs). This in
turn allows to use standard learning algorithms (e.g. SVM) to solve the
predictive task.
